Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-02-02 WSSSPE3/report/WSSSPE3_review.tex
Section 3: touch up.

WSSSPE3 left the traditional Supercomputing Conference environment this year,
and in Turk's words, this spoke to the fact that scientific software comes from
many different types of inquiries, deployment, strategies for maintenance,
users, and ways of measuring the value of a piece of software. It appeared to
Turk that the supercomputing community generally adopts some top-down
approaches, whereas the SciPy community more often than not uses more bottom-up
systems. The essential messages perceived were also often bipolar: the
supercomputing community thinks that software is getting harder, with exascale
computing and optimization issues in mind; but the SciPy community thinks that
software is becoming better, with emerging tools such as Jupyter and
productivity packages for research workflows. Admitting such comparisons are
somewhat unfair generalizations, Turk reminded the audience that the different
approaches bring different types of ideas to the table, and he welcomed WSSSPE3
being conducted outside existing preconceptions.

Returning to the topic of his talk, Turk invited the audience to picture
scientific software as a flower on a landscape under the Sun, which may
represent a number of measurable factors such as number of citations; growth of
a community and number of contributors; amount of funding; prestigious prizes
awarded; stability of the community in terms of leadership transitions, serving
community needs, not breaking test suites, and performance on new architectures.
But all these metrics are strictly speaking \emph{proxies} for the values and
the impact scientific software bears. What we can measure does not give us
direct insight---it just gives us proxies of insight.

@@

Turk said that sustaining scientific software really meant to him conducting
scientific inquiries, often by some specific software, and sustaining the people
we care about, our careers, and the future of our fields. According to Turk, we
all have an invested stake in sustaining scientific software. Hence, having
``sustained'' projects can suffocate new projects, so we need to make sure we
don't cause novel ideas and packages to suffer at the hands of the status quo.

@@

\item How can productized software balance its future versus its past, or the
new needs of the customers versus the existing needs of the development
community?

>>>>

WSSSPE3 left the traditional Supercomputing Conference environment this year,
and in Turk's words, this change spoke to the fact that scientific software comes from
many different types of inquiries, deployment, strategies for maintenance,
users, and ways of measuring the value of a piece of software. It appeared to
Turk that the supercomputing community generally adopts some top-down
approaches, whereas the SciPy community more often than not uses more bottom-up
systems. The essential messages perceived were also often bipolar: the
supercomputing community thinks that software is getting harder, with exascale
computing and optimization issues in mind; but the SciPy community thinks that
software is becoming better, with emerging tools such as Jupyter and
productivity packages for research workflows. Admitting such comparisons are
somewhat unfair generalizations, Turk reminded the audience that the different
approaches bring different types of ideas to the table, and he welcomed WSSSPE3
being conducted outside existing preconceptions.

Returning to the topic of his talk, Turk invited the audience to picture
scientific software as a flower on a landscape under the Sun, which may
represent a number of measurable factors such as number of citations; growth of
a community; number of contributors; amount of funding; prestigious prizes
awarded; stability of the community in terms of leadership transitions; serving
community needs; not breaking test suites; and performance on new architectures.
But all these metrics are strictly speaking \emph{proxies} for the values and
the impact scientific software bears. What we can measure does not give us
direct insight---it just gives us proxies of insight.

@@

Turk said that sustaining scientific software really meant to him conducting
scientific inquiries, often by some specific software, and sustaining the people
we care about, our careers, and the future of our fields. According to Turk, we
all have an invested stake in sustaining scientific software. Hence, having
``sustained'' projects can suffocate new projects, so we need to make sure we
do not cause novel ideas and packages to suffer at the hands of the status quo.

@@

\item How can productized software balance its future versus its past, or the
emerging needs of the customers versus the existing needs of the development
community?
